### 2021-10-04 preregistration/preregistration_analyses.Rmd
reformat task folders

---
title: "preregistration_analyses"
author: "Stefan Vermeent"
date: "9/29/2021"
output: pdf_document
---

```{r libraries, include=FALSE}
library(tidyverse)
```