### 2021-10-18 preregistration/preregistration_analyses.Rmd
intermediate update prereg

---
bibliography: references.bib
csl: apa.csl
output:
  pdf_document: default
  officedown::rdocx_document:
    reference_docx: reference-doc.docx
    page_margins:
      bottom: 1
      top: 1
      right: 1
      left: 1
      header: 0.5
      footer: 0
      gutter: 0
    plots:
      align: center
      caption:
        style: Image Caption
        pre: 'Figure '
        sep: '. '
    tables:
      caption:
        style: Table Caption
        pre: 'Table '
        sep: '. '
---

```{r libraries, include=FALSE}
library(tidyverse)
```

# Methods {#methods}

# Participants {#participants}

Participants will be XXX US-based individuals between the ages of 18 and 30. We decided for this age cut-off to limit the effect of healthy age-related cognitive decline on speeded cognitive tasks, which becomes more pronounced after age 30 (@salthouse_selective_2010).

# Procedure {#procedure}

## Measures {#measures}

This is the measures section.

### Attention tasks {#tasks}

This is the section featuring all the attention tasks.

### Questionnaires {#questionnaires}

This is the section featuring all the questionnaires.

## Data analyses {#analyses}

lorem ipsum

### preprocessing {#preprocessing}

lorem ipsum

### Reaction Times {#rt}

lorem ipsum

### Drift Diffusion Models {#ddm}

lorem ipsum