### 2022-02-07 preregistrations/1_pilot/scripts/old/3_DDM_estimation.Rmd

---
title: "Drift Diffusion Model"
bibliography: ../references.bib
---

```{r, echo=T, results='hide', message=F, warning=F} 
library(tidyverse)
library(here)
library(RWiener)
library(mvtnorm)
library(flankr)

source(here("preregistrations", "1_pilot", "scripts", "custom_functions", "functions_exclusions.R"))

change_data_clean <- read_csv(here("data", "1_pilot", "change_data_clean.csv")) 
cueing_data_clean <- read_csv(here("data", "1_pilot", "cueing_data_clean.csv"))
flanker_data_clean <- read_csv(here("data", "1_pilot", "flanker_data_clean.csv")) 
```

# Change Detection Task

## Data Pre-treatment

First, we check whether there is a significant difference between 'same' and 'different' conditions on the Change Detection Task.

```{r}
change_DDM_setup <- change_data_clean %>%
  select(id, rt, correct, condition) %>%
  mutate(rt = rt / 1000)

change_DDM_setup %>%
  group_by(id, condition) %>%
  summarise(mean_rt = mean(log(rt), na.rm=T)) %>%
  ungroup() %>%
  pivot_wider(names_from = "condition", values_from = "mean_rt") %>%
  do(broom::tidy(t.test(.$same, .$different, data=., paired=T)))
```

```{r}
for (i in unique(change_DDM_setup$id)) {
  # 1. Filter subject i and write individual data files to data folder
  change_DDM_i <- change_DDM_setup %>%
    filter(id == i) %>%
    select(-id, -condition) %>%
    write_delim(file = here("data", "1_pilot", "DDM", str_c("change_DDM_subject", i, ".dat")), col_names = FALSE)
}

# 2. Run DDM model
path_to_files <- paste0("pushd ", here("data", "1_pilot", "DDM")) %>% str_replace_all(., pattern="/", replacement="\\\\")
run_DDM <- " && fast-dm.exe change_ml.ctl"
cmd <- paste0(path_to_files, run_DDM)

shell(cmd)

# 3. Read results
change_DDM_results <- read_delim(here("data", "1_pilot", "DDM", "ddm_results_change.lst"), delim = " ", trim_ws = TRUE) %>%
  mutate(dataset = str_replace_all(dataset, "^change_DDM_subject", "") %>% as.numeric(),
         task = "change") %>%
  rename(id = dataset)

# 4. Remove data files from folder
files_to_remove <- list.files(here("data", "1_pilot", "DDM"), pattern = ".dat$|.lst$", full.names=TRUE)
file.remove(files_to_remove)

change_DDM_results

change_DDM_results %>%
  pivot_longer(c(a,v,t0), names_to = "parameter", values_to = "value") %>%
  ggplot(aes(value)) +
  geom_histogram() +
  theme_classic() +
  facet_wrap(~parameter, scales = "free")
```

## Assessing Model Fit

```{r}
# Means
mu_v     <- mean(change_DDM_results$v, na.rm=T)
mu_a     <- mean(change_DDM_results$a, na.rm=T)
mu_t0    <- mean(change_DDM_results$t0, na.rm=T)

# SDs
sd_v     <- sd(change_DDM_results$v, na.rm=T)
sd_a     <- sd(change_DDM_results$a, na.rm=T)
sd_t0    <- sd(change_DDM_results$t0, na.rm=T)

# Correlations
cor_v_a  <- cor(change_DDM_results$v, change_DDM_results$a, use="complete.obs")
cor_v_t0 <- cor(change_DDM_results$v, change_DDM_results$t0, use="complete.obs")
cor_a_t0 <- cor(change_DDM_results$a, change_DDM_results$t0, use="complete.obs")

# Variance-Covariance Matrix
sigma <- matrix(c(sd_v^2, sd_v*sd_a*cor_v_a, sd_v*sd_t0*cor_v_t0,
                  sd_a*sd_v*cor_v_a, sd_a^2, sd_a*sd_t0*cor_a_t0,
                  sd_t0*sd_v*cor_v_t0, sd_t0*sd_a*cor_a_t0, sd_t0^2),
                ncol=3)

nsim <- 5000

# Simulate DDM parameters based on empirical data
sim_params <- rmvnorm(n=nsim, mean = c(mu_v, mu_a, mu_t0), sigma = sigma) %>%
  as_tibble() %>%
  rename(v=V1, a=V2, t0=V3) %>%
  mutate(n=50, z = 0.5, row = 1:n()) %>%
  select(n, a, t0, z, v, row) %>%
  filter(a > 0 & t0 > 0)

if(nrow(sim_params) < nsim) warning(paste0(nsim - nrow(sim_params < nsim), " simulated driftrates or non-decision times were smaller than 0 and have been dropped!"))
```

```{r echo = FALSE, results = FALSE}
# Simulate and save RTs
pmap(sim_params, function(n,a,t0,z,v, row) {
  rwiener(n,a,t0,z,v) %>%
    as_tibble() %>%
    mutate(resp = ifelse(resp == "upper", 1, 0)) %>%
    write_delim(file = here("data", "1_pilot", "DDM", str_c("par_sim", row, ".dat")), col_names = FALSE)

})
```

```{r}
# Run DDM model on simulated datasets
path_to_files <- paste0("pushd ", here("data", "1_pilot", "DDM")) %>% str_replace_all(., pattern="/", replacement="\\\\")
run_DDM <- " && fast-dm.exe change_ml.ctl"
cmd <- paste0(path_to_files, run_DDM)

shell(cmd)

# Read results
fit_vector <- read_delim(here("data", "1_pilot", "DDM", "ddm_results_change.lst"), delim = " ", trim_ws = TRUE) %>%
  select(fit) %>%
  pull()

# Remove data files from folder
files_to_remove <- list.files(here("data", "1_pilot", "DDM"), pattern = ".dat$|.lst$", full.names=TRUE)
file.remove(files_to_remove)

# Determine the 5% quantile of the distribution of fit indices to use as critical value
fit_critical_value <- quantile(fit_vector, .05)[[1]]
```

```{r}
change_DDM_results %<>%
  mutate(
    fit_critical_value = fit_critical_value,
    ex_change_bad_fit = ifelse(fit < fit_critical_value, TRUE, FALSE))

```

Using `r nsim` simulations, we found that the 5% quantile of the fit distribution corresponded to a critical value of `r round(fit_critical_value,2)`. Applying this critical value to the obtained fit statistics in the data of the Change Detection Task, we found that the data of `r sum(change_DDM_results$ex_change_bad_fit)` subjects fell below this cut-off.

# Attention Cueing Task

For the Attention Cueing Task, we have two conditions: the neutral condition and the cued condition. DDM parameters can be fixed across conditions or set to vary across conditions. It seems plausible that the manipulation across trials has an effect on the non-decision time. For the neutral trial, subjects are only able to make a saccade in the correct direction when the target is presented. Therefore, there is an offset in the time at which information can be accumulated relative the the cued trial. Second, the drift rate might be affected by the manipulation

## Model 1. Fully independent: All parameters vary across conditions

```{r}
cueing_DDM_setup <- cueing_data_clean %>%
  select(id, rt, correct, condition) %>%
  mutate(rt = rt / 1000)

for (i in unique(cueing_DDM_setup$id)) {
  # 1. Filter subject i and write individual data files to data folder
  cueing_DDM_i <- cueing_DDM_setup %>%
    filter(id == i) %>%
    select(-id) %>%
    write_delim(file = here("data", "1_pilot", "DDM", str_c("cueing_DDM_subject", i, ".dat")), col_names = FALSE)
}

# 2. Run DDM model
path_to_files <- paste0("pushd ", here("data", "1_pilot", "DDM")) %>% str_replace_all(., pattern="/", replacement="\\\\")
run_DDM <- " && fast-dm.exe cueing_full_independent_ml.ctl"
cmd <- paste0(path_to_files, run_DDM)

shell(cmd)

# 3. Read results
cueing_DDM_results1 <- read_delim(here("data", "1_pilot", "DDM", "ddm_results_cueing_full_independent.lst"), delim = " ", trim_ws = TRUE) %>%
  mutate(dataset = str_replace_all(dataset, "^cueing_DDM_subject", "") %>% as.numeric(),
         iteration  = "full_independent",
         task = "cueing") %>%
  rename(id = dataset)

# 4. Remove data files from folder
files_to_remove <- list.files(here("data", "1_pilot", "DDM"), pattern = ".dat$|.lst$", full.names=TRUE)
file.remove(files_to_remove)
```

## Model 2. Boundary separation is fixed across conditions

```{r}
for (i in unique(cueing_DDM_setup$id)) {
  # 1. Filter subject i and write individual data files to data folder
  cueing_DDM_i <- cueing_DDM_setup %>%
    filter(id == i) %>%
    select(-id) %>%
    write_delim(file = here("data", "1_pilot", "DDM", str_c("cueing_DDM_subject", i, ".dat")), col_names = FALSE)
}

# 2. Run DDM model
path_to_files <- paste0("pushd ", here("data", "1_pilot", "DDM")) %>% str_replace_all(., pattern="/", replacement="\\\\")
run_DDM <- " && fast-dm.exe cueing_v_t0_independent_ml.ctl"
cmd <- paste0(path_to_files, run_DDM)

shell(cmd)

# 3. Read results
cueing_DDM_results2 <- read_delim(here("data", "1_pilot", "DDM", "ddm_results_cueing_v_t0_independent.lst"), delim = " ", trim_ws = TRUE) %>%
  mutate(dataset = str_replace_all(dataset, "^cueing_DDM_subject", "") %>% as.numeric(),
         iteration = "a_independent",
         task = "cueing",
         a_neutral = a,
         a_cued = a) %>%
  rename(id = dataset) %>%
  select(-a)

# 4. Remove data files from folder
files_to_remove <- list.files(here("data", "1_pilot", "DDM"), pattern = ".dat$|.lst$", full.names=TRUE)
file.remove(files_to_remove)

cueing_DDM_results2

save(change_DDM_results, cueing_DDM_results1, cueing_DDM_results2, file = here("data", "1_pilot", "DDM", "DDM_objects.RData"))
```

# Comparisons between models

```{r}
# T-test on boundary separation
t.test(cueing_DDM_results1$a_cued, cueing_DDM_results1$a_neutral, paired = T)

bind_rows(cueing_DDM_results1, cueing_DDM_results2) %>%
  pivot_longer(ends_with(c("neutral", "cued")), names_to = "parameter", values_to = "value") %>%
  select(id, iteration, parameter, value) %>%
  pivot_wider(names_from = "iteration", values_from = "value") %>%
  ggplot(aes(full_independent, a_independent)) +
  geom_point() +
  facet_wrap(~parameter, scales = "free") +
  labs(
    title = "Parameter comparison across model versions"
  )

bind_rows(cueing_DDM_results1, cueing_DDM_results2) %>%
  select(id, iteration, fit) %>%
  pivot_wider(names_from = "iteration", values_from = "fit") %>%
  ggplot(aes(full_independent, a_independent)) +
  geom_point() +
  labs(
    title = "Model fit comparison across model versions"
  )

cueing_DDM_results1 %>% select(v_cued, t0_cued, a_cued, v_neutral, t0_neutral, a_neutral) %>% cor %>% corrplot(method = "number")
cueing_DDM_results2 %>% select(ends_with(c("cued", "neutral"))) %>% cor %>% corrplot(method = "number")

```

```{r}
cueing_DDM_results1 %>%
  pivot_longer(ends_with(c("neutral", "cued")), names_to = "parameter", values_to = "value") %>%
  separate(parameter, c("parameter", "condition"), sep = "_") %>%
  pivot_wider(names_from = "condition", values_from = "value") %>%
  mutate(
    parameter = case_when(
      parameter == "v" ~ "Drift rate",
      parameter == "t0" ~ "Non-decision time",
      parameter == "a" ~ "Boundary separation"
    )
  ) %>%
  ggplot(aes(neutral, cued)) +
  geom_point() +
  theme_classic() +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~parameter, scales = "free") +
  labs(
    title = "Model 1: Boundary separation varies across conditions"
  )

cueing_DDM_results2 %>%
  pivot_longer(ends_with(c("neutral", "cued")), names_to = "parameter", values_to = "value") %>%
  separate(parameter, c("parameter", "condition"), sep = "_") %>%
  pivot_wider(names_from = "condition", values_from = "value") %>%
  mutate(
    parameter = case_when(
      parameter == "v" ~ "Drift rate",
      parameter == "t0" ~ "Non-decision time",
      parameter == "a" ~ "Boundary separation"
    )
  ) %>%
  ggplot(aes(neutral, cued)) +
  geom_point() +
  theme_classic() +
  geom_abline(slope = 1, intercept = 0) +
  facet_wrap(~parameter, scales = "free") +
  labs(
    title = "Model 2: Boundary separation is fixed across conditions"
  )
```
## Assessing Model Fit

We assess Model fit using Monte Carlo simulations based on the empirical data, as detailed in [@voss_2015].

```{r}
# Means
mu_v_neut     <- mean(cueing_DDM_results2$v_neutral, na.rm=T)
mu_a_neut     <- mean(cueing_DDM_results2$a_neutral, na.rm=T)
mu_t0_neut    <- mean(cueing_DDM_results2$t0_neutral, na.rm=T)

mu_v_cued        <- mean(cueing_DDM_results2$v_cued, na.rm=T)
mu_a_cued        <- mean(cueing_DDM_results2$a_cued, na.rm=T)
mu_t0_cued       <- mean(cueing_DDM_results2$t0_cued, na.rm=T)

# Initiate Variance-Covariance Matrix
sigma <- matrix(nrow = 6, ncol = 6, dimnames=list(c("sd_v_neutral", "sd_a_neutral", "sd_t0_neutral", "sd_v_cued", "sd_a_cued", "sd_t0_cued"),
                                         c("sd_v_neutral", "sd_a_neutral", "sd_t0_neutral", "sd_v_cued", "sd_a_cued", "sd_t0_cued")))

# Fill cells of Var-Covar matrix 
for (col in 1:ncol(sigma)) {
  for (row in 1:nrow(sigma)) {

    col_i = colnames(sigma)[col]
    row_i = rownames(sigma)[row]

    # Variances
    if(col_i == row_i) {
      param <- str_replace(col_i, "^sd_", "")

      sigma[col_i, row_i] <- sd(cueing_DDM_results2[[param]], na.rm = T)^2
    } else {

      # Covariances
      param_col <- str_replace(col_i, "^sd_", "")
      param_row <- str_replace(row_i, "^sd_", "")

      sd_col <- sd(cueing_DDM_results2[[param_col]], na.rm = T)
      sd_row <- sd(cueing_DDM_results2[[param_row]], na.rm = T)
      r      <- cor(cueing_DDM_results2[[param_col]], cueing_DDM_results2[[param_row]])

      sigma[col_i, row_i] <- sd_col * sd_row * r
    }
  }
}

nsim <- 5000

# Simulate DDM parameters based on empirical data
sim_params <- rmvnorm(n=nsim, mean = c(mu_v_neut, mu_a_neut, mu_t0_neut, mu_v_cued, mu_a_cued, mu_t0_cued), sigma = sigma) %>%
  as_tibble() %>%
  rename(v_neut=V1, a_neut=V2, t0_neut=V3, v_cued=V4, a_cued=V5, t0_cued=V6) %>%
  mutate(n=64, z = 0.5, row = 1:n()) %>%
  select(n, a_neut, t0_neut, z, v_neut, a_cued, t0_cued, z, v_cued, row) %>%
  filter(across(starts_with(c("a_", "t0_")), ~ . > 0))

if(nrow(sim_params) < nsim) warning(paste0(nsim - nrow(sim_params < nsim), " simulated driftrates or non-decision times were smaller than 0 and have been dropped!"))
```

```{r echo = FALSE, results = FALSE}
# Simulate and save RTs
pmap(sim_params[c("n", "a_neut", "t0_neut", "z", "v_neut", "row")], function(n,a_neut,t0_neut,z,v_neut, row) {
  rwiener(n,a_neut,t0_neut,z,v_neut) %>%
    as_tibble() %>%
    mutate(resp = ifelse(resp == "upper", 1, 0)) #%>%
    #write_delim(file = here("data", "1_pilot", "DDM", str_c("par_sim", row, ".dat")), col_names = FALSE)
})

```

```{r}
cueing_DDM_results2
change_DDM_results
```

# Flanker Task

```{r}
flanker_DDM_setup <- flanker_data_clean %>%
  select(subject=id, congruency, accuracy=correct, rt) %>%
  mutate(rt = rt / 1000) 
```

```{r}
flanker_fit <- fitSSP(flanker_DDM_setup)
```

```{r}
#write_csv(change_DDM_results, here("data", "1_pilot", "DDM_results_change.csv"))
#write_csv(flanker_DDM_results, here("data", "1_pilot", "DDM_results_flanker.csv"))
#write_csv(cueing_DDM_results, here("data", "1_pilot", "DDM_results_cueing.csv"))
```