### 2023-06-06 preregistrations/3_study2/preregistration_amendment.Rmd
amendment to preregistration for study 2

---
title: "Amendment to preregistration"
output: github_document
---

Go to [this link](https://github.com/stefanvermeent/attention_project/tree/main/preregistrations/3_study2) for the original preregistration.

## Amendment to preregistration for study 2 on `r Sys.Date()`

### Global-Local performance

In the original preregistration, we specified that we would exclude participants who performed at chance on either the Flanker Task or the Global-Local Task, which was defined as an accuracy of 59.4% or lower.
However, initial inspections of the Global-Local Task data showed that a substantial part of the sample did not reach this cut-off, suggesting that the task was more difficult than anticipated.
Thus, we developed a more fine-grained approach (described below) in an attempt to distinguish between 1) participants who did not understand the task and 2) participants who understood the task, but found it difficult to perform well. 
Given the assumptions of DDM, the first group would have to be excluded because they likely did not go through a process of information accumulation.
However, the model should be able to adequately fit the data of the second group.

The amended analysis approach for the Global-Local Task looks as follows:

1) Fit the data to the cleaned data of the full sample, including participants who performed at or below chance level (i.e., after trial-level exclusions but before case-wise exclusions);
2) Based on recovered parameter estimates for each participant, simulate the same number of trials (reaction times and accuracy) using the `RWiener` package, separately for Global and Local trials.
3) For each participant, calculate the 25th, 50th and 75th quantile of both their real RTs and the simulated RTs. 
In addition, we calculate mean accuracies based on the real and simulated data.
4. Compute standardized residuals between the real and simulated data for RTs at each quantile and for accuracy.
In case of good fit, the residual should be close to zero.
5. Exclude the data of participants with any standardized residual > 3.2 SD.

### Multiverse analyses

In the preregistration, we planned to include three variables as covariates that were previously featured as arbitrary exclusion decisions in the multiverse specification: 1) whether or not participants rescaled the screen; 2) whether or not participants exited fullscreen mode at some point during the tasks; 3) Whether or not participants experienced interruptions during the tasks.
Our reasoning was that these three factors were consistently found to have a large impact on model results.
However, we realized later on that adding these factors as covariates was not a good approach from a causal inference standpoint.
That is, it is more likely that each of these factors added random noise to our estimates than that they had a causal effect on the outcome.
Therefore, we decided instead to include these factors as arbitrary decisions in the multiverse analyses, similarly as the previous two studies.
This will allow for a coherent assessment of influential factors across all three experiments.