### 2024-03-28 manuscript/titlepage.qmd
anonymize GitHub links

---
bibliography: references.bib
csl: apa.csl
format:
  docx:
    reference-doc: reference-doc.docx
output:
  officedown::rdocx_document:
    page_margins:
      bottom: 1
      footer: 0
      gutter: 0
      header: 0.5
      left: 1
      right: 1
      top: 1
    plots:
      align: center
      caption:
        pre: 'Figure '
        sep: '. '
        style: Image Caption
    tables:
      caption:
        pre: 'Table '
        sep: '. '
        style: Table Caption
  pdf_document: default
  word_document: default
editor: 
  markdown: 
    wrap: sentence
---

#### **Childhood adversity is not associated with lowered inhibition, but slower perceptual processing: A Drift Diffusion Model analysis**

#### Stefan Vermeent^a,b^, Ethan S. Young^a^, Jean-Louis van Gelder^b^, & Willem E. Frankenhuis^a,b,c^

#### ^a^ Department of Psychology, Utrecht University, Utrecht, The Netherlands

#### ^b^ Max Planck Institute for the Study of Crime, Security, and Law, Freiburg, Germany

#### ^c^ Evolutionary and Population Biology, Institute for Biodiversity and Ecosystem Dynamics, University of Amsterdam, Amsterdam, the Netherlands

<br>

# Corresponding author

Correspondence concerning this article should be addressed to Stefan Vermeent, Department of Psychology, Utrecht University, Heidelberglaan 1, 3584 CS Utrecht, The Netherlands. Email: p.c.s.vermeent@gmail.com / p.c.s.vermeent@uu.nl.

# Data Statement

All scripts, materials, and data supporting the results presented here are available on the article's GitHub repository [ANONYMIZED FOR REVIEW].

# Funding Statement
WEF’s contributions have been supported by the Dutch Research Council (V1.Vidi.195.130) and the James S. McDonnell Foundation (https://doi.org/10.37717/220020502).

# Author contributions

**Stefan Vermeent:** Conceptualization, Data curation, Formal analysis, Investigation, Methodology, Project administration, Data curation, Visualization, Writing - original draft, Writing - review & editing.
**Ethan S. Young:** Conceptualization, Methodology, Supervision, Writing - review & editing.
**Jean-Louis van Gelder:** Conceptualization, Funding acquisition, Methodology, Supervision, Writing - review & editing.
**Willem E. Frankenhuis:** Conceptualization, Funding acquisition, Methodology, Supervision, Writing - review & editing.