### 2024-05-28 preregistrations/README.Rmd
README file for preregistrations folder

---
title: ""
output: 
  html_document:
    template: ../assets/template.html
    css: ../assets/style.css
editor_options: 
  markdown: 
    wrap: sentence
---

# Links to analyses scripts

1.  [Pilot study](https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/1_pilot/scripts)
2.  [Study 1](https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/2_study1/scripts)
3.  [Study 2](https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/3_study2/scripts)

# Links to Cognitive task materials

1.  [Pilot study](https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/1_pilot/materials)
2.  [Study 1](https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/2_study1/materials)
3.  [Study 2](https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/3_study2/materials)

# Links to preregistrations

1.  [Pilot study](https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/1_pilot)
2.  [Study 1](https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/2_study1)
3.  [Study 2](https://github.com/StefanVermeent/attention_project/tree/main/preregistrations/3_study2)