### 2024-05-31 manuscript/CD_submission1/abstract.qmd
cleaning up manuscript folder

---
bibliography: ../references.bib
csl: apa.csl
format:
  docx:
    reference-doc: ../reference-doc.docx
output:
  officedown::rdocx_document:
    page_margins:
      bottom: 1
      footer: 0
      gutter: 0
      header: 0.5
      left: 1
      right: 1
      top: 1
    plots:
      align: center
      caption:
        pre: 'Figure '
        sep: '. '
        style: Image Caption
    tables:
      caption:
        pre: 'Table '
        sep: '. '
        style: Table Caption
  pdf_document: default
  word_document: default
editor: 
  markdown: 
    wrap: sentence
---

# Abstract

It is well-established that individuals who grew up in adverse conditions tend to be slower on the Flanker Task. 
This finding is typically interpreted to reflect difficulty inhibiting distractions. 
However, it might result from slower general cognitive processes (e.g., reduced general processing speed), rather than the specific ability of inhibition. 
We used Drift Diffusion Modeling in three online studies (total N = 1560) with young adults to understand associations of adversity with Flanker performance. 
We find no associations between exposure to violence and unpredictability with inhibition. 
Yet, although mixed, violence and unpredictability exposure were associated with lower strength of perceptual input--—how well someone can process target and distractor information alike. 
Finally, people with lower strength of perceptual input processed information more holistically, focusing less on details. 
Thus, lowered Flanker performance does not necessarily imply lowered inhibition ability. 
Cognitive modeling reveals a different picture of abilities in adverse conditions as opposed to analyses based on raw performance.

*Key-words:* childhood adversity, inhibition, Flanker, Drift Diffusion Modeling